New theoretical lower bounds of weak Schur numbers using regular Schur numbers

### 2021-02-23 demonstration_romain.tex

Schur proved in [1] that given a number of subsets $n$, there exists a value of $m$ such that there exists no partition of $[\![1,p]\!]$ for any $p \geqslant m$. This observation leads to the following definition.

@@

The values of both Schur and weak Schur numbers are known only for small values of $n$, and our goal is to improve the lower bounds on weak Schur numbers for $n \geqslant 5$. At the time being, the best lower bounds for weak Schur numbers are established by Rowley in [2]. He greatly improved the former best values with a purely theoretical and constructive approach. More precisely, he managed to build interesting weakly sum-free partitions with sum-free partitions. With this, he found the following lower bounds :

@@

- If $u \in [\![n + 1,\beta - 1]\!]$ then $x = (a+1)\beta - (\beta - u)$ with $-\alpha \leqslant \beta - u \leqslant 0$ and so $x \in \mathcal{B}$ \\

@@

>>>>

Schur proved in \hyperlink{label1}{\textbf{[1]}} that given a number of subsets $n$, there exists a value of $m$ such that there exists no partition of $[\![1,p]\!]$ in $n$ sum-free subsets for any $p \geqslant m$. This observation leads to the following definition.

@@

The values of both Schur and weak Schur numbers are known only for small values of $n$, and our goal is to improve the lower bounds on weak Schur numbers for $n \geqslant 5$. At the time being, the best lower bounds for weak Schur numbers are established by Rowley in \hyperlink{label2}{\textbf{[2]}}. He greatly improved the former best values with a purely theoretical and constructive approach. More precisely, he managed to build interesting weakly sum-free partitions with sum-free partitions. With this, he found the following lower bounds :

@@

- If $u \in [\![n + 1,\beta - 1]\!]$ then $x = (a+1)\beta - (\beta - u)$ with $-\alpha \leqslant \beta - u \leqslant 0$. Furthermore, $a \leqslant p - 1$, else we would have $x > N$, and so $x \in \mathcal{B}$ \\

@@

This first partition of $[\![1,N]\!]$ will help us to define our final partition by the projection of its equivalence relation. We thereby define $h : [\![1,N]\!] \longrightarrow [\![1,r+s]\!]$ as such :\\
- If $x \in \mathcal{A}$ then $h(x) = f(x \text{ mod } \beta)$ (well defined because $x \text{ mod } \beta \in [\![1,N]\!]$)\\
- If $x \in \mathcal{B}$ then $x = a\beta + u$ with a unique $(a,u) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]$ and we define $h(x) = r + g(a)$\\
The fact that $(\mathcal{A}, \mathcal{B})$ is a partition of $[\![1,N]\!]$ ensures that this definition of $h$ is valid. We then have to verify that $h$ induces weakly sum-free subsets.\\
\underline{The classes of equivalence $h(x)$ for $x \in \mathcal{A}$ are weakly sum-free :}
Let $(x,y) \in \mathcal{A}^2$ such that $h(x) = h(y)$, $x \neq y$ and $x + y \leqslant N$
\item If $(x,y) \in [\![1,\alpha]\!]^2$ :\\
We have $x + y \leqslant 2\alpha \leqslant n$ and $x + y = 0\beta + x + y$, therefore $x + y \in \mathcal{A}$. Then, by definition : $h(x) = f(x)$, $h(y) = f(y)$ and $h(x+y) = f(x+y)$, which gives us, thanks to the property verified by $f$, that $h(x+y) \neq h(x)$.
\item If $(x,y) \in [\![1,\alpha]\!] \times ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] )$ :\\
We write $y = a\beta + u$ with $(a,u) \in [\![0,p]\!] \times [\![\alpha + 1,n]\!]$. Then $x+y = a\beta + x + u = (a+1)\beta + x + u - \beta$, and if $x + u > n$ it follows that $a \leqslant p-1$ since $x+y \leqslant N$, and $-\alpha \leqslant x + u - \beta \leqslant -1$. Therefore $x+y \in \mathcal{B}$ and $h(x+y) \neq h(x) = f(x)$ by definition of h. On the contrary, if $x - u \leqslant n$, then $x+y \in \mathcal{A}$ and $h(x+y) = f(x+u)$ because $x+u$ is actually the remainder of the euclidean division of $x+y$ by $\beta$. Moreover, $h(x) = f(x)$, $x < u$ and, with our initial hypothesis, $h(x) = h(y) = f(u)$. The property verified by $f$ gives us $f(x+u) \neq f(x)$ which can be rewritten as $h(x+y) \neq h(x)$.
\item If $(x,y) \in ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] ) \times [\![1,\alpha]\!]$ : \\
This case is handled exactly like the previous one by swaping the roles of $x$ and $y$.
\item If $(x,y) \in ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] )^2$ : \\
We write $x = a\beta + u$ and $y = b\beta + v$ with $(a,u)$ and $(b,v)$ in $[\![0,p]\!] \times [\![\alpha + 1,n]\!]$. Then $x+y = (a+b)\beta + u+v = (a+b+1)\beta + u + v - \beta$ with $a+b \leqslant p-1$ (else we would have $x+y > N$ because $u+v > n$) and $-\alpha \leqslant u + v - \beta \leqslant \alpha$, therefore $x+y \in \mathcal{B}$ and by definition $h(x+y) \neq h(x)$.
In any case, $h(x+y) \neq h(x)$ and the classes of equivalence $h(x)$ for $x \in \mathcal{A}$ are weakly sum-free.\\
\underline{The classes of equivalence $h(x)$ for $x \in \mathcal{B}$ are weakly sum-free :}
Let $(x,y) \in \mathcal{B}^2$ such that $h(x) = h(y)$, $x \neq y$ and $x + y \leqslant N$.\\
We write $x = a\beta + u$ and $y = b\beta + v$ with $(a,u)$ and $(b,v)$ in $[\![1,p]\!] \times [\![-\alpha,\alpha]\!]$. We have $h(x) = r + g(a)$ and $h(y) = r + g(b)$, therefore $g(a) = g(b)$. We also have $x+y = (a+b)\beta + u +v$.\\
If $u + v \in [\![-\alpha,\alpha]\!]$, then $x+y \in \mathcal{B}$ and $h(x+y) = g(a+b)$, hence we can deduce that $h(x+y) \neq h(x)$ because of the property verified by $g$. On the contrary, if $u+v \notin [\![-\alpha,\alpha]\!]$, then necessarily $x+y \in \mathcal{A}$. Suppose $x+y \in \mathcal{B}$, then $x+y = c\beta + w$ with $(c,w) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]$. Thus, $c\beta + w = (a+b)\beta + u + v$ and $(a+b-c)\beta = w-u-v$. Furthermore $a+b-c \neq 0$, else we would have $u+v = w \in [\![-\alpha,\alpha]\!]$. This finally leads to the following inequality :
\[\beta \leqslant |a+b-c|\beta = |w-u-v| \leqslant |w| + |u| + |v| \leqslant 3\alpha \leqslant n + \alpha < \beta
which is absurd. We can therefore conclude that $x+y \in \mathcal{A}$ and by definition of $h$, $h(x+y) \neq h(x)$, proving that the classes of equivalence $h(x)$ for $x \in \mathcal{B}$ are weakly sum-free.\\
Finally, we have showed that every classe of equivalence induced by $h$ is weakly sum-free, which ends the proof.
\section{References}
\hypertarget{label1}{\textbf{[1]}} Schur, I.. "Über Kongruenz x ... (mod. p.).." Jahresbericht der Deutschen Mathematiker-Vereinigung 25 (1917): 114-116. \url{http://eudml.org/doc/145475} \\
\hypertarget{label2}{\textbf{[2]}} F. Rowley, New Lower Bounds for Weak Schur Partitions, arXiv 2011.11292, \url{https://arxiv.org/pdf/2011.11292.pdf}
